---
title: "`r paste0(snakemake@params$genus, ' — primer comparison across genes')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_depth: 2
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
suppressPackageStartupMessages({
  library(knitr)
  library(kableExtra)
})

genus       <- snakemake@params$genus
genes       <- snakemake@params$genes
primers_in  <- snakemake@input$primers
pcr_in      <- snakemake@input$pcr
```

## Overview

- **Genus:** `r genus`
- **Genes compared:** `r paste(genes, collapse = ", ")`
- **Generated:** `r format(Sys.time(), "%Y-%m-%d %H:%M %Z")`

```{r assemble}
read_safe <- function(path) {
  tryCatch(read.delim(path, stringsAsFactors = FALSE),
           error = function(e) data.frame())
}

# Per-gene summary stats
summary_rows <- list()
top_pair_rows <- list()
for (i in seq_along(genes)) {
  g       <- genes[i]
  primers <- read_safe(primers_in[i])
  pcr     <- read_safe(pcr_in[i])

  if (nrow(primers) == 0) {
    summary_rows[[g]] <- data.frame(
      gene = g, n_pairs = 0,
      best_score = NA, best_retrieval = NA,
      mean_amplicon_len = NA
    )
    next
  }

  best <- primers[1, ]
  best_pcr <- if (nrow(pcr) > 0) pcr[pcr$primer_id == best$primer_id, ] else NULL
  best_retrieval <- if (!is.null(best_pcr) && nrow(best_pcr) > 0)
    as.numeric(best_pcr$retrieval_rate) else NA
  mean_amp <- if (!is.null(best_pcr) && nrow(best_pcr) > 0)
    as.numeric(best_pcr$mean_amplicon_len) else NA

  summary_rows[[g]] <- data.frame(
    gene = g,
    n_pairs = nrow(primers),
    best_score = round(best$score, 3),
    best_retrieval = best_retrieval,
    mean_amplicon_len = mean_amp
  )

  best$gene <- g
  top_pair_rows[[g]] <- best
}

summary_df <- do.call(rbind, summary_rows)
```

## Per-gene summary

```{r summary, results='asis'}
disp <- summary_df
disp$best_retrieval <- ifelse(
  is.na(disp$best_retrieval),
  "—",
  sprintf("%.1f%%", 100 * disp$best_retrieval)
)
disp$mean_amplicon_len <- ifelse(
  is.na(disp$mean_amplicon_len), "—",
  sprintf("%.0f bp", disp$mean_amplicon_len)
)

kable(disp, format = "html", row.names = FALSE,
      col.names = c("Gene", "# pairs", "Best score",
                    "Best-pair retrieval", "Mean amplicon")) |>
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE)
```

## Best primer pair per gene

```{r best, results='asis'}
if (length(top_pair_rows) == 0) {
  cat("No primer pairs available for any gene.")
} else {
  top_df <- do.call(rbind, top_pair_rows)
  show <- top_df[, c("gene", "primer_id", "fwd", "rev",
                     "amplicon_len", "fwd_tm", "rev_tm",
                     "fwd_deg", "rev_deg", "score")]
  kable(show, format = "html", row.names = FALSE) |>
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE, font_size = 12)
}
```

## How to use this comparison

When choosing a marker gene for amplicon sequencing, weigh:

- **Number of valid primer pairs** — more options = more flexibility for
  optimization
- **Best-pair retrieval rate** — fraction of target-genus genomes that
  amplify; high values mean broad coverage
- **Mean amplicon length** — should fit your sequencing platform's read length
- **Score** — combines primer-site conservation, internal-region diversity,
  and Tm matching, so higher is better for taxonomic resolution

The "right" gene depends on biological context (lifestyle conservation,
copy number, sequencing platform), so use these numbers as one input to
the decision, not a verdict.
