---
title: "`r paste0(snakemake@params$genus, ' / ', snakemake@wildcards$gene, ' primer report')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_depth: 2
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

suppressPackageStartupMessages({
  library(knitr)
  library(kableExtra)
})

genus       <- snakemake@params$genus
gene        <- snakemake@wildcards$gene
primers_tsv <- snakemake@input$primers
pcr_tsv     <- snakemake@input$pcr
plot_path   <- snakemake@input$plot
```

## Summary

- **Genus:** `r genus`
- **Gene:** `r gene`
- **Generated:** `r format(Sys.time(), "%Y-%m-%d %H:%M %Z")`

```{r load}
primers <- tryCatch(
  read.delim(primers_tsv, stringsAsFactors = FALSE),
  error = function(e) data.frame()
)
pcr <- tryCatch(
  read.delim(pcr_tsv, stringsAsFactors = FALSE),
  error = function(e) data.frame()
)
```

## Top primer pairs

```{r top, results='asis'}
if (nrow(primers) == 0) {
  cat("**No primer pairs passed the filters.** Try relaxing the GC, Tm, ",
      "or amplicon-length range in `config.yaml`.\n", sep = "")
} else {
  show <- primers[, c("primer_id", "fwd", "rev",
                      "amplicon_len", "fwd_tm", "rev_tm", "tm_diff",
                      "fwd_gc", "rev_gc",
                      "fwd_deg", "rev_deg", "score")]
  show$fwd_gc <- sprintf("%.1f%%", 100 * show$fwd_gc)
  show$rev_gc <- sprintf("%.1f%%", 100 * show$rev_gc)
  kable(show, format = "html", row.names = FALSE,
        caption = sprintf("Top %d primer pairs (ranked by combined score)",
                          nrow(show))) |>
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = FALSE, font_size = 12)
}
```

## In silico PCR validation

```{r pcr, results='asis'}
if (nrow(pcr) == 0) {
  cat("No in silico PCR results available.\n")
} else {
  pcr$retrieval_rate <- sprintf("%.1f%%", 100 * as.numeric(pcr$retrieval_rate))
  kable(pcr, format = "html", row.names = FALSE,
        caption = "Per-pair retrieval rate against all genomes") |>
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = FALSE, font_size = 12)
}
```

## Diversity profile

```{r div, out.width="100%"}
if (file.exists(plot_path)) {
  knitr::include_graphics(plot_path)
} else {
  cat("(diversity plot not generated)")
}
```

## Notes

- **Score interpretation:** higher is better. The score combines primer-site
  conservation, internal-region diversity, and Tm matching.
- **Degeneracy fold:** the number of distinct primer sequences represented by
  the IUPAC code. A fold of 1 means a fully specified primer; higher folds
  give broader coverage at the cost of synthesis complexity.
- **Retrieval rate** is the fraction of genomes from which `seqkit amplicon`
  successfully extracts a product (allowing up to 3 mismatches per primer).
